---
title: "Data Processing Hands On Practice"
output: pdf_document
---

## Setup

Load necessary package(s) and import the ```income1``` and ```income2``` datasets into R. The ```income1``` dataset contains demographics information of 30162 individuals, and the ```income2``` dataset contains income information for the same set of individuals.
```{r}
library(dplyr)
income1 = read.csv("income1.csv")
income2 = read.csv("income2.csv")
income = left_join(income1, income2, by='id')
```


## Answer the following questions. You need to decide which dataset(s) to use.

Q1. How many male and female individuals are there in the dataset, and what is the average age for each gender group?
```{r}
income %>% group_by(sex) %>%
  summarise(n = n(),
            ave_age = mean(age))
```


Q2. What are the maximum and minimum working hours per **day** (assuming a 5-day work week) for each occupation?
```{r}
income %>% mutate(hours.per.day = hours.per.week/5) %>% 
  group_by(occupation) %>%
  summarise(max_hpd = max(hours.per.day),
            min_hpd = min(hours.per.day))

```


Q3. Find the average working hours per week for each education level, then rank the results by average working hours from long to short.
```{r}
income %>% group_by(education) %>%
  summarise(ave_hpw = mean(hours.per.week)) %>% 
  arrange(desc(ave_hpw))

```

Q4. How many people, whose wages are higher than 50K, are there in each work class?
```{r}
income %>% group_by(workclass) %>%
  filter(wage == '>50K') %>%
  summarise(n = n())

```


Q5. Among males who are older than 40, what proportion makes less than 50K?
```{r}


```